---
title: "PML-Project: Classification based on a set of variables to predict the manner in which the excersie is done"
author: "Mahmood Karimi"
date: "27 Aug 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
Based on these data we are going to build a prediction model, which predicts the manner in which the excersize is done according to quantified variables.

## Data and environment preparation
```{r warning=FALSE, message=FALSE}
library(caret)
library(ggplot2)
library(randomForest)
library(corrplot)
library(rpart)
set.seed(1000)

mainDir <- getwd()
subDir <- "outputDirectory"
if (!file.exists(subDir)){
        dir.create(file.path(mainDir, subDir))
}

if(!file.exists("outputDirectory/training.csv")){
        fileURL_train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
        download.file(fileURL_train, destfile = "outputDirectory/training.csv")
        
}

if(!file.exists("outputDirectory/test.csv")){
        fileURL_test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
        download.file(fileURL_test, destfile = "outputDirectory/test.csv")
        
}

pmltr <- read.csv("outputDirectory/training.csv", stringsAsFactors = FALSE)
pmlts <- read.csv("outputDirectory/test.csv", stringsAsFactors = FALSE)

# remove index variable which does not affect on prediction
pmltr <- pmltr[, -c(1, 3:7)]
pmltr$user_name <- as.factor(pmltr$user_name)
pmltr$classe <- as.factor(pmltr$classe)

# convert variables to number
for(i in 1:ncol(pmltr)){
        
        if(class(pmltr[,i]) == "character"){
                
                pmltr[,i] <- as.numeric(pmltr[,i])
        }
}
# remove variables which contraint NA
num_na <- sapply(1:ncol(pmltr), function(x) sum(is.na(pmltr[,x])))
pmltr_new <- pmltr[, num_na == 0]

```

## Explorarory analysis
```{r eda}
table(pmltr_new$classe)
table(pmltr_new$user_name)
# draw some plot of data
qplot(user_name, roll_belt, colour=classe, data=pmltr_new, main = "Activity level of each user on belt", xlab = "user name", ylab = "activity level on belt")
```

## Partition training data
```{r part}
inTrain <- createDataPartition(y = pmltr_new$classe, p = 0.75, list = FALSE)
tempTrain <- pmltr_new[inTrain,]
tempTest <- pmltr_new[-inTrain,]
table(tempTrain$classe)
```

## Create Models: Classification Trees, Random Forest, Gradient Boosting Model
```{r models}
ct_model <- train(classe ~., data = tempTrain, method = "rpart", trControl=trainControl(method="cv",number=5))
#rf_model <- train(classe ~., data = tempTrain, method = "rf")
#gbm_model <- train(classe ~., data = tempTrain, method = "gbm", verbose = FALSE)
```

## Display outputs
```{r output}
confusionMatrix(predict(ct_model, newdata = tempTest), tempTest$classe)
print(plot(varImp(ct_model, scale = FALSE)))
#confusionMatrix(predict(rf_model, newdata = tempTest), tempTest$classe)
#confusionMatrix(predict(gbm_model, newdata = tempTest), tempTest$classe)

predict(ct_model, newdata = pmlts)
#predict(rf_model, newdata = pmlts)
#predict(gbm_model, newdata = pmlts)
```


# End of Report

